Visualization and analysis of probability distributions of large temporal data

### 2022-02-28 report/response.Rmd

> Q4. On page 35, refer to months consistently using short or long names.

The comment is addressed by changing the x-axis labels to month-of-year for Figures 3.1 (page 36) and 3.2 (page 38).

> Q5. On page 35 v refers to the number of variables, in Section 3.3.3 it refers to the variable.

On p38, $v$ changed to $p$, where $p$ represents the number of variables.

Clarification: This is analogous to many time granularities and their pairwise combinations and $v$ represents the univariate measured variable for which distributions are constructed for different time granularities and their pairs (added to Table 3.1 on page 42).

### 2022-02-28 report/response.Rmd

Added "for a continuous univariate dependent variable" in the last paragraph of section 3.1 (Introduction) on page 39.

> Q3. In Figure 3.1, the points are a bit hard to see, so it looks like whiskers extend beyond the data.

The following are changed in Figure 3.1 to address the comment on page 36:

 - changed to theme_bw() instead of default  
 - outliers and jitters are made more transparent  
 - jitter width reduced to make it clear that whiskers do not extend beyond data

@@

> Q21. In Figure 3.6 the axis tick labels should be smaller. The blue and orange are hard to see. The caption
should refer to the rug. Maybe show fewer panels?

Changed (page 52). Axis tick labels made smaller and the rug plots are removed to make the display less busy.

> Q22. In Figure 3.7 (a) the heatmaps need id labels. The grey color is missing from the legend. Maybe use a different colour in the heatmap for the significant comparisons. State the threshold for significance in the caption. In Table 3.3 the caption should explain the threshold. (Maybe use colour instead of stars to indicate significance?)

The following changes have been made to Figure 3.7 (a) (page 53):

 - different box borders (instead of id labels) are added such that (a) and (b) display the household ids in the same order as indicated by the same colour of the line plot in (b) and box border
 - grey color explained in the text and caption

@@

Text added under Section 4.2.2 on pages 64 and 65 to clarify this.

<!-- Clarification: Each observation (time series) is scaled independently in both JS-NQT and JS-RS. For the WPD approach, each observation might be scaled or not scaled, resulting in usable groups in either case. -->

> Q7. Page 64 “D is the Jensen-Shannon distance”.

I have mentioned that $D$ stands for Jensen-Shannon distances (now page 66).

@@

The table format and caption changed to address this comment. (page 70)

<!-- No change. Stating the distribution of each $v_t$ is difficult because the levels for each granularity and design varies, therefore I chose to divide it into two tables. -->

> Q10. Where is $\mu$ in Table 4.2?

The changed caption (Table 4.2 page 70) will help to clarify this.

@@

> Q5. Before introducing a new distance measure in Section 3.2, I would like to see a literature review of existing distance measures and their limitations. Why is it necessary to introduce a new distance measure here? This is partially addressed on p. 41, but that should be placed earlier in the text. One clarifying question: Which density is represented by f in the equation? And shouldn't there be a dx at the end of the integral? Also, the distance measures that are mentioned at the end of that paragraph should be explained in one or two sentences.

Changed the following:

- Introduction (last but one paragraph on page 39) to incorporate comments on existing distance measures and their limitations.

- "Distance between distributions" (under section 3.2.3 on page 44) to mention why we chose Jensen-Shannon distances and some details about other distance measures.

### 2022-02-28 report/response.Rmd

<!-- Clarification: Null distributions refer to the case when there is no significant distributional differences between categories of one or more granularities. Data in Fig 3.3 (a) is consistent with the null distribution. -->

> Q10. In table 3.1 $N_c$ is the number of cyclic granularities, whereas in the text (first sentence of 3.3.2 and
3.3.3) it is m.

Added $m$ to table 3.1 (page 42).
$m$: number of cyclic granularities to display together

Clarification: $N_c$ refers to the total number of contextual cyclic granularities. $m$ refers to the number of cyclic granularities we are considering together in the display. For example, contextual cyclic granularities could be hour_day, day_week and month_year and we want to visualize any one granularity at a time. So $N_c = 3$ and $m=1$.

@@

Done.

> Q5. Before introducing a new distance measure in Section 3.2, I would like to see a literature review of existing distance measures and their limitations. Why is it necessary to introduce a new distance measure here? This is partially addressed on p. 41, but that should be placed earlier in the text. One clarifying question: Which density is represented by f in the equation? And shouldn't there be a dx at the end of the integral? Also, the distance measures that are mentioned at the end of that paragraph should be explained in one or two sentences.

Changed the following:

- Introduction (last but one paragraph on page 39) to incorporate comments on existing distance measures and their limitations

- "Distance between distributions" (under section 3.2.3 on page 44) to mention why we chose Jensen-Shannon distances and some details about other distance measures

- f in the equation was a typo. Replaced by log now

- dx added at the end of the integral

@@

> Q17. In Figure 3.7a, it is not clear what the eight heatmaps represent. Do they belong to id 1, ..., id 8 ? Some explanation in the text or listing the id information on the left would be helpful. Similarly, on p. 53, the text states: id 7 and 8 have the same significant harmonies. I do not know where to look at the heatmaps in Figure 3.7a to find this information (and what is mentioned under items 2. and 3. in the text). Some explanation and labeling is needed here.

Caption changed on page 53 to illustrate the plot better. Added box borders in (a) such that (a) and (b) display the household ids in the same order as indicated by the same color of the line plot in (b) and box border in (a). Added text on page 55 to state where to look at for item 2 and 3 in the text.

> Q18. How were the harmony pairs selected and sorted in Figure 3.7 and Table 3.3? headmaps show 25 pairs, the parallel coordinate plot shows only 14 (same as Table 3.3). However, the last two are sorted differently, "hod wdwnd" (the first row in the table) is neither the first nor the last row in the parallel coordinate plot.

Changed the text description of Figure 3.7 (page 55) and Table 3.3 (page 54) to address the comments.

@@

Caption rephrased to make it more clear on page 56.

<!-- For id: 7, there is more distributional difference between consecutive hours of the day compared to id:1. Also all the days of the week are different from each other in id:7. id:1 has similar behavior across all days within both weekdays and weekends. For this reason, the wpd for id:7 has a higher value than id:1 as the similar behavior within weekdays and weekends balances out the difference between weekdays and weekends. The display with hod vs wdwnd would have shown higher values of wpd for id:1 (as can be verified from Table 3). -->

> Q22. I didn't ask earlier, but how were the eight households for Section 3.4 being selected? Was this based on a random sample, or manually do obtain rather different energy usage patterns? This should be answered earlier on, but this also should be addressed in the Discussion: What can be expected when working with the full data set of Chapter 2, i.e., about 13,000. You indicate:
A future direction of work is to be able to explore and compare many individuals/subjects together for similar patterns across significant granularities. What would be computationally (and visually) feasible as 13,000 is a few orders of magnitude
higher than the number of eight households used in this example? Clearly, no full answer is expected here, but rather some speculation of what could possibly be done in the future (or what is done in the next chapter).

Text added in the first paragraph of Section 3.4 to elaborate how households were selected (page 51).

A paragraph is added to the Discussion section to incorporate comments on computational time (page 57).

### 2022-02-28 Rmd/03-hakear.Rmd
correcting minor inconsistencies - all done

Let $v_{i, t}$ denote the electricity demand for the $i^{th}$ household in time period $t$. The series $\{v_{i, 1},\dots,v_{i,T}\}$ is the linear granularity corresponding to half-hour since the interval of the `tsibble` is 30 minutes. We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar. From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow* and *dom*  respectively. Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior (abbreviated as *wnwd*). Thus, seven cyclic granularities are considered to be of interest. The pairs of cyclic granularities ($C_{N_C}$) will have $7\times 6=42$ elements. The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in @Gupta2021-gravitas. Table \ref{tab:tab-rank-8} shows $14$ harmony pairs that belong to $H_{N_C}$.

@@

$\wpd_{i}$ is computed on $v_{i, t}$ for all harmony pairs $\in H_{N_C}$ and for each household $i \in \{1, 2, \dots, 8\}$. The harmony pairs are then arranged in descending order and highlighted with `***`, `**` and `*` corresponding to the $99^{th}$, $95^{th}$ and $90^{th}$ percentile threshold. Table \ref{tab:tab-rank-8} shows the rank of the harmonies for different households. The rankings are different for different households, which is a reflection of their varied behaviors. Most importantly, there are at most three harmonies that are significant for any household. This is a huge reduction in the number of potential harmonies to explore.

@@

Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs. Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities. The colors (in shades of red) represent the value of $\wpd$ for each of the harmony pairs (in Table \ref{tab:tab-rank-8}) and the gray tiles correspond to clashes. A darker shade of red corresponds to higher values of $\wpd$. Those with `*` correspond to $\wpd$ values above $\wpdsub{threshold95}$.

>>>>

Let $v_{i, t}$ denote the electricity demand for the $i^{th}$ household in time period $t$. The series $\{v_{i, 1},\dots,v_{i,T}\}$ is the linear granularity corresponding to half-hour since the interval of the `tsibble` is 30 minutes. We consider coarser linear granularities like hour, day, week and month from the commonly used Gregorian calendar. From the four linear granularities of hour, day, week, and month, we obtain $N_C = 4\times 3/2 = 6$ cyclic granularities:  "hour_day", "hour_week", "hour_month", "day_ week", "day_month" and "week_month" abbreviated as *hod*, *how*, *hom*, *dow*, *dom* and *wom* respectively. Further, we add cyclic granularity day-type ("wknd wday") to capture weekend and weekday behavior (abbreviated as *wnwd*). Thus, seven cyclic granularities are considered to be of interest. The pairs of cyclic granularities ($C_{N_C}$) will have $7\times 6=42$ elements. The set of possible harmonies $H_{N_C}$ from $C_{N_C}$ are chosen by removing clashes using procedures described in @Gupta2021-gravitas. Table \ref{tab:tab-rank-8} shows $14$ harmony pairs that belong to $H_{N_C}$.

@@

$\wpd_{i}$ is computed on $v_{i, t}$ for all harmony pairs $\in H_{N_C}$ and for each household $i \in \{1, 2, \dots, 8\}$. The harmony pairs are then arranged in descending order and highlighted with `***`, `**` and `*` corresponding to the $99^{th}$, $95^{th}$ and $90^{th}$ percentile threshold. Table \ref{tab:tab-rank-8} shows the rank of the harmonies for different households and the harmonies are sorted according to id 1. The rankings are different for different households, which is a reflection of their varied behaviors. Most importantly, there are at most three harmonies that are significant for any household. This is a huge reduction in the number of potential harmonies to explore.

@@

Figure \ref{fig:dotplot-8} helps to compare households through the heatmap (a) across harmony pairs (lexicographic order). Each household is represented by 25 tiles, each tile representing a pair of cyclic granularities with x-variables on the horizontal axis and facet-variables on the vertical axis. The colors (in shades of red) represent the value of $\wpd$ for each of the harmony pairs (in Table \ref{tab:tab-rank-8}) and the gray tiles correspond to clashes. A darker shade of red corresponds to higher values of $\wpd$. Those with `*` correspond to $\wpd$ values above $\wpdsub{threshold95}$.
